A Multi-Modal Model of Object Deformation under Robotic Pushing

### 2016-12-24 response-to-reviewers-second-resubmission.txt

While the authors recognize the issue of the on-line learning in the introduction (page 1, col 2, line 35), they do not address that in the paper.

From your comments, we believe that mentioning on-line learning, while important, was misleading in terms of our paper, so we have withdrawn mention of it.

Also they refer the problem of getting access to partial views of the objects but it seems that in this paper the object is viewed from a convenient point of view.

We have added text to make clear the restriction in our study of using a single view chosen by the user.

@@

What happen if we add a new material? Do we need to add a new feature space?

If the object is plastic or elastic then we don’t need to add a new feature space, but we would be wise to calibrate a new model. We have added a comment to this effect in the conclusion.

I would add a figure like Fig. 9a at this point. In this way the reader can have an idea of the setup, how vision acquire information and how the mass-spring model depends on the visual information itself. For example, it might be interesting to combine Fig 9a and Fig 2 in single sketch of the model.

Thank you for this suggestion. We have prepared a new version of Figure 2, as you suggest. This incorporates elements from the old Figure 2 and Figure 9.

A consideration about the organisation of the following sections. Given that the framework consists of two models, force prediction and shape prediction, I would expect section III.1 to be Force prediction and III.2 to the shape prediction (or vice-versa). However, the organization of Sections IV, V, VI, in my opinion is misleading.

We have reorganised the paper as you request.

Finally, I would start describing the algorithm from the visual system. It seems that the visual system has a very important role in this paper but it is not highlighted. Indeed, the mass-spring model described is section IV should be created within the perimeter of the segmented object. Thus, it requires a visual system capable of localizing and segmenting the object. In my understanding, this information is necessary to initialize the position of the particles (described in section IV) inside the object. Regarding this point, I would like to know how the particles are placed inside the object. What happens if we change the size of the object? Do we need to change the number of particles? What happen if the object is not rectangular?

Eq. 19. I really do not understand how the force could be proportional to the velocity. This point should be clarified. My understanding is that the authors assumed the acceleration (a) to be constant during the interval h, so that h*F = h*a*m => F = v*m/h. Given that h is a constant, the force (F), the velocity (v) or the mass (m) are recovered with a scale factor. Is that correct? If so, what is the difference between this approach and the quasi-static approach proposed by Frank. Please, explain.

This section is quite long and I think it could be simplified notably. The authors could simply state the
energy E = Sum(i=1:N) [ 0.5 k_i * C_i(p1, pn)^2], where N is the number of constraints. There are several constraints to preserve the length, the area, etc. and the force is given as negative gradient of the potential
energy with respect to the position of the particle. Details about all the equation could be reported in an appendix to make the paper more readable. About the equation, personally I don’t like the symbol composed of multiple letters such as Area, forcemag, forcedir, etc.

We have shortened the section as requested.

@@

This section is quite long and I think it could be simplified notably. The authors could simply state the
energy E = Sum(i=1:N) [ 0.5 k_i * C_i(p1, pn)^2], where N is the number of constraints. There are several constraints to preserve the length, the area, etc. and the force is given as negative gradient of the potential
energy with respect to the position of the particle. Details about all the equation could be reported in an appendix to make the paper more readable. About the equation, personally I don’t like the symbol composed of multiple letters such as Area, forcemag, forcedir, etc.

We have shortened the section as requested.

Do the authors use just one video for the training and two video for the testing? It seems that the dataset is quite small to have a systematic test of the algorithm. Even if there are only two objects, the authors should test different motions of the finger (i.e. pushing velocity, penetration, …)

This is rather difficult to modify as we gathered the data set three years ago, and it would not be possible to extend it quickly enough using the same set up. However, to try to satisfy your request we are currently working on an extension of the data set using simulated data from a physics engine. While this will not be physically accurate it will allow further testing of the model. We are afraid we have not yet completed this work for this revision, as we focussed on completing the study of the comparable method, but we will submit our additional data and the results with a new revision when they are available.

@@

I don’t see the results of the classification. I would expect a confusion matrix to show how many times the sponge was classified as sponge and as plasticine and vice-versa. Table III seems to show the results about the classification of the pixels.

Vero?

Moreover, I would like to see a comparison against a baseline approach. They could some of the models proposed in the literature or some simple models. For example, the authors could assume that the material “disappears” when the finger penetrates the object and report the F-Score of this method. Alternatively, they could use a simple model with only one spring between the finger and the base of the object.

We have implemented a growing neural gas network previously published by Cretu et al. We have added additional results in a comparison of this method with our own.

@@

Page 12. Col 2. Line 3. Please, can you clarify the hand tuning? It seems that the genetic algorithm is not the best way to find the parameters? Have you tried to use other method like gradient descent? Is it possible to use the model to mass-spring model to predict the force exerted on the finger? In that case, the authors could use this prediction error as fitness function and use the gradient descent to find the parameters.

Vero?

Other comments:

Fig. 1 is not referenced in the paper.

We have decided that this figure added little, hence it was not referenced naturally anywhere in the text. We have thus decided to remove it to save space, since other requested changes have added considerable length.

Page 2, Col 1, Line 17. Section VI is not mentioned.

Page 2. Col 2, Line 5: FEM is not defined. Finite-element models?

We have defined this.

Fig 2. Instead of use symbols, I would write Vision, strain, force, etc.

Thank you, this chimes with the comments of Reviewer 2, we have fixed this in Figures 1, 2 and 3.

Page 3, Col 1, line 57. Cretu et al. -> cite [10]?

We have added this reference.

Page 8. Col 2, Line 27. Linear snake = linear spline?

Algorithm 5. I would specify that the F-Score is the fitness function of
the algorithm. Moreover, it would be nice to know the values used for the
parameters of the algorithm (perhaps using a table).
Page 10. Col 1. Line 45. Please specify that there are two objects, a
sponge and plasticine, with a rectangular shape and so on.

### 2016-12-25 twostagedo.tex
Answered to most technical details.  Two are missing.

When one or more particles are displaced from their equilibirum positions, the forces that tend to restore an elastic body to its original shape are modeled by the forces of springs (along the edges of the mesh) that try to recover their original length).  These forces act upon the particles modifying their positions in time.  Traditionally, the relationship between the position of the particle $p_i$ and those forces $F$ obeys the second law of Newton:
\begin{align}
 F = m \dfrac{\partial^2 p}{{\partial t}^2} \label{eq:newton}
\end{align}
where $t$ is time.  However, it is possible to deviate from this equation if we need to model a different behaviour. %are interested in is numerical accuracy, as it will be done here.

@@

To simulate the effect of an external actuator (like the finger of the robot) interacting with the deformable object, external forces $F_{ext}$ acting on vertices of the mesh must be added to \eref{eq:newton}:
\begin{align}
 F + F_{ext} = m \dfrac{\partial^2 p}{{\partial t}^2}
\end{align}
Another option could be to induce response forces by displacing vertices from their position of equilibrium.
}

@@

The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004}, which is explained below, was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.  The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.  Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.  The potential energy terms for the preservation of length with damping and preservation of area were included in our system \comment{exactly as they were defined by Teschner}.  To compensate for the lack of a term for preservation of volume, which tends to prevent distortions of the graph \comment{in the original model}, \comment{we derived } a new term for the preservation of the angles \comment{in every triangle} \sout{was included}.  The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.

\comment{Teschner defined a generalized version of the simple mass-spring model by using the concept of \emph{constraint}.  He makes use of different constraints to model the tendency of an object to recover several of its attributes.} For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.  When the value of the constraint is zero, the potential energy is zero as well.  When the constraint is not satisfied, the potential energy increases.  This behaviour is obtained through the following definition for the energy:
\begin{align}
 E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint}
\end{align}
where $k$ is a proportionality constant \comment{\sout{\eref{eq:EConstraint}}} and $p_i$ is the position of particle $i$.

In order to reduce the energy of the system when a constraint is not satisfied, the force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
\begin{align}
 \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
\end{align}
Given the oscillatory nature of these equations, \comment{when using constraints that model the behaviour of springs,} damping can be introduced to stop the oscillations.  This is modelled with a force acting in the opposite direction to the velocity of the affected particles\comment{, thus slowing down motion.  Teschner wrote this as a function of the constraints as well \cite{Teschner2004}}:
\begin{align}
 \label{eq:Fdamping}
 \vec{F}^i&(\vec{p}_1,...,\vec{p}_{n},\vec{v}_1,...,\vec{v}_{n}) = \nonumber \\
    &\left( -kC-k_d\sum_{1 \le j \le n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
 \dfrac{\partial C}{\partial \vec{p}_i},
\end{align}
where $\vec{v}_j = \frac{\partial \vec{p}_j}{\partial t}$ is the velocity of the $j^{th}$ particle connected to $i$, when the constraint acts on $n$ particles.

@@

\subsubsection{Preservation of Area}
This constraint is applied per triangle in the mesh.  Since Teschner and Morris \comment{%\sout{didn't find it helpful to add damping to the preservation of areas, it is avoided here.} 
found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.}  Therefore the force $\vec{F}^i_A$ is derived as follows:
\begin{align}
 C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0} \\
 Area &= \frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|, \\
 \vec{F}^i_A(p_i,p_j,p_k) &= -\frac{k_A}{2A_0^2} \left( Area-A_0 \right) \nonumber \\
 & \frac{(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)}{|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|} [\vec{1} \times (\vec{p}_j-\vec{p}_k)] \label{eq:area_force}
\end{align}
where $A_0$ is the initial area of the triangle and $k_A$ the proportionality constant.

@@

\subsubsection{Preservation of Angles}
The previous terms can not do anything to restore the mesh if triangles get flipped during deformation.  For this reason we included a new term that enforces the preservation of the original angles of the triangles.  [\fref{fig:forces}(c)] gives an idea of how these forces look.  Here the energy depends on the difference \comment{%\sout{of the angles between adjacent edges.} 
between the current angle, between adjacent edges, and the angle between them at the equilibrium position.} Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
\begin{align}
 E_\varphi(\varphi) & = \frac{1}{2}k_\varphi(\varphi - \varphi_0)^2 \\
 \varphi(p_i,p_j,p_k) & = \arccos \left( \dfrac{(p_j - p_i) \cdot (p_k - p_i)}{ \left\| p_j - p_i \right\| \left\|p_k - p_i \right\|} \right)^2 \nonumber
\end{align}

@@

\subsection{Plastic Deformation}
\begin{figure}
 \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
 %\STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
 \comment{\STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.} And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
 \IF {$\alpha > max\_\alpha$}
  \STATE $l_0 \gets l$
 \ELSE
  \STATE $l_0 \gets (1.0 - \alpha) D_0$
 \ENDIF
 \STATE $elastic\_deformation \gets \frac{l_0 - l}{D_0}$
 \IF {$elastic\_deformation > yield$}
  \STATE $\alpha \gets \alpha + creep * elastic\_deformation$
  \IF {$\alpha > max\_\alpha$}
   \STATE $\alpha \gets max\_\alpha$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
\caption{Algorithm to calculate the permanent deformation of an edge.}
\end{figure}

@@

\subsection{Integration Scheme}
\label{sec:ischeme}
\comment{Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equiations,} we made the \comment{force proportional to the velocity:
\begin{align}
 F(t) = m \dfrac{\partial \vec{p}}{\partial t}
\end{align}
and applied Euler only once to solve for $p$:
\begin{equation}
 \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
\end{equation}}
with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.  \comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is no inertia in the oscillation.
When experimenting with different integration steps, force values were interpolated between measurements.

\subsection{Collision Detection}
In the simulation, rigid objects \comment{(like the robotic finger and the table)} are treated as geometric obstacles.  \comment{When the simulation of the mesh deformation runs---used to model the sponge---the predicted positions of the vertices may indicate that the mesh penetrates an obstacle.  This should be interpreted as a collision between the sponge and an obstacle.  Since the sponge can not penetrate an obstacle, these overlaps must be eliminated. %The physical consequences of this will affect the system predictions.  For example: if the vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.
}

\comment{In order to model this, we model the finger and table using geometric primitives.  For the table obstacle, any interpenetration by the sponge is solved by pushing vertices of the mesh of the sponge back to the border of the table.  The springs will then propagate the effect to their neighbours.  For the finger obstacle, resolution is in two stages: by vertex and then by edge.}  A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.  For edges, the \comment{standard} algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.  Given this distance, the algorithm in \alref{alg:push_edge} indicates how far and in which direction the edge must be displaced.

@@

\subsection{\comment{Additional} Geometric Constraints}
\comment{While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function. Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.  We achieve this through geometric checking subroutines.}  There are two geometric constraints that the mesh is enforced to maintain:

\begin{enumerate}
 \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed more than half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ of this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.  This is enough for most cases, but triangles in the border of the mesh can still overlap each other.

@@

The next stage \comment{is} the learning process \comment{which} consists in identifying sets of parameters that produce simulations similar to the ground truth.  This problem is addressed in the following section.

\section{\comment{Learning to Predict Deformations}}
\label{sec:learning}

@@

The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. \comment{Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.}  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \comment{\cite{Menet1991}. The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.  Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy. This is explained below.}  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  \comment{  Since the object of interest has a distinctive colour, the hue of the pixels on every edge are used to keep only edges that could belong to it.  If the average hue of the pixels on an edge falls outside a previously calibrated range, the edge is eliminated.}
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains where it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

@@

\subsection{Evolutionary Algorithm}
\comment{Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they instead depend on the geometry of the mesh used to model the object. There can thus be several sets of parameters in different regions of the parameter space that produce adequate simulations \cite{Morris2008}. We therefore chose to use an evolutionary algorithm to both perform global search, and local improvement.}

The evolutionary algorithm described in \alref{alg:genetic} was used to search the parameter space of the mass-spring model presented in the previous section.
From the parameters that must be chosen for the simulation to work, some were fixed while others had to be searched for.  They are:
\begin{itemize}
 \item The mass per vertex.  \textit{(For the moment we consider a unitary mass for all vertices.)}
 \item The elasticity constants: for the preservation of length $k_L$, area $k_A$, angles $k_{\varphi}$ and linear damping $k_d$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The plastic parameters: $yield$, $creep$ and the maximum amount of permanent deformation $max\_\alpha$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The integration time step $\Delta t$.  \textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well.)}
 \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
 \item The minimum area before a triangle is considered \textit{flat}.  (It has been set to $2.0$ units.)
\end{itemize}
To evaluate the suitability of each set of parameters, the simulation is compared with the ground truth tracked from real objects.

>>>>

When one or more particles are displaced from their equilibirum positions, the forces that tend to restore an elastic body to its original shape are modeled by the forces of springs (along the edges of the mesh) that try to recover their original length).  These forces act upon the particles modifying their positions in time.  Traditionally, the relationship between the position of the particle $p_i$ and those forces $F$ obeys the second law of Newton:
\begin{align}
 F = m \dfrac{\partial \vec{p}}{\partial t} \label{eq:newton}
\end{align}
where $t$ is time.  However, it is possible to deviate from this equation if we need to model a different behaviour. %are interested in is numerical accuracy, as it will be done here.

@@

To simulate the effect of an external actuator (like the finger of the robot) interacting with the deformable object, external forces $F_{ext}$ acting on vertices of the mesh must be added to \eref{eq:newton}:
\begin{align}
 F + F_{ext} = m \dfrac{\partial \vec{p}}{\partial t}
\end{align}
Another option could be to induce response forces by displacing vertices from their position of equilibrium.

@@

The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004}, which is explained below, was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.  The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.  Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.  \comment{Since the location of the masses and direction of the springs will bias the deformation response of the object, it is important to distribute the masses uniformly and place the springs with a symetric arrangement for an isotropic material.  This placement can be done with an algorithm that takes into account the symmetries of the object at the begining of the experiment.  If the material gets permanently deformed a properly callibrated physics based model, like the one we propose here, will produce an adequate new arrangement from which new interactions can be studied.}

The potential energy terms for the preservation of length with damping and preservation of area were included in our system exactly as they were defined by Teschner.  To compensate for the lack of a term for preservation of volume, which tends to prevent distortions of the graph in the original model, we derived a new term for the preservation of the angles in every triangle.  The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.

Teschner defined a generalized version of the simple mass-spring model by using the concept of \emph{constraint}.  He makes use of different constraints to model the tendency of an object to recover several of its attributes. For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.  When the value of the constraint is zero, the potential energy is zero as well.  When the constraint is not satisfied, the potential energy increases.  This behaviour is obtained through the following definition for the energy:
\begin{align}
 E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint}
\end{align}
where $k$ is a proportionality constant and $p_i$ is the position of particle $i$.

In order to reduce the energy of the system when a constraint is not satisfied, the force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
\begin{align}
 \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
\end{align}
Given the oscillatory nature of these equations, when using constraints that model the behaviour of springs, damping can be introduced to stop the oscillations.  This is modelled with a force acting in the opposite direction to the velocity of the affected particles, thus slowing down motion.  Teschner wrote this as a function of the constraints as well \cite{Teschner2004}:
\begin{align}
 \label{eq:Fdamping}
 \vec{F}^i&(\vec{p}_1,...,\vec{p}_{n},\vec{v}_1,...,\vec{v}_{n}) = \nonumber \\
    &\left( -kC-k_d\sum_{1 \le j \le n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
 \dfrac{\partial C}{\partial \vec{p}_i},
\end{align}
where $\vec{v}_j = \frac{\partial \vec{p}_j}{\partial t}$ is the velocity of the $j^{th}$ particle connected to $i$, when the constraint acts on $n$ particles.

@@

\subsubsection{Preservation of Area}
This constraint is applied per triangle in the mesh.  Since Teschner and Morris
found that the use of damping for the preservation of areas does not significantly improve the stability of the simulation, it is not included here either.  Therefore the force $\vec{F}^i_A$ is derived as follows:
\begin{align}
 C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0} \\
 Area &= \frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|, \\
 \vec{F}^i_A(p_i,p_j,p_k) &= -\frac{k_A}{2A_0^2} \left( Area-A_0 \right) \nonumber \\
 & \frac{(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)}{|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|} [\vec{1} \times (\vec{p}_j-\vec{p}_k)] \label{eq:area_force}
\end{align}
where $A_0$ is the initial area of the triangle and $k_A$ the proportionality constant.

@@

\subsubsection{Preservation of Angles}
The previous terms can not do anything to restore the mesh if triangles get flipped during deformation.  For this reason we included a new term that enforces the preservation of the original angles of the triangles.  [\fref{fig:forces}(c)] gives an idea of how these forces look.  Here the energy depends on the difference between the current angle, between adjacent edges, and the angle between them at the equilibrium position. Energy is given by\footnote{It was also considered to multiply $E_\varphi$ by the lengths of the edges, but it hasn't improved the performance of the model.}:
\begin{align}
 E_\varphi(\varphi) & = \frac{1}{2}k_\varphi(\varphi - \varphi_0)^2 \\
 \varphi(p_i,p_j,p_k) & = \arccos \left( \dfrac{(p_j - p_i) \cdot (p_k - p_i)}{ \left\| p_j - p_i \right\| \left\|p_k - p_i \right\|} \right)^2 \nonumber
\end{align}

@@

\subsection{Plastic Deformation}
\begin{figure}
 \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
 \STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
 \IF {$\alpha > max\_\alpha$}
  \STATE $l_0 \gets l$
 \ELSE
  \STATE $l_0 \gets (1.0 - \alpha) D_0$
 \ENDIF
 \STATE $elastic\_deformation \gets \frac{l_0 - l}{D_0}$
 \IF {$elastic\_deformation > yield$}
  \STATE $\alpha \gets \alpha + creep * elastic\_deformation$
  \IF {$\alpha > max\_\alpha$}
   \STATE $\alpha \gets max\_\alpha$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
\caption{Algorithm to calculate the permanent deformation of an edge.}
\end{figure}

@@

\subsection{Integration Scheme}
\label{sec:ischeme}
Rather than using traditional Newtonian equations with a numeric integration scheme like Euler or Verlet, used twice to solve the second order differential equiations, we made the force proportional to the velocity:
\comment{
\begin{align}
 F(t) = \vec{p}
\end{align}
which implies the following assumptions:
\begin{align}
 \dfrac{\partial \vec{p}}{\partial t} = \vec{p}
\end{align}
this equation can only be satisfied with an exponential, thus avoiding the oscillatory behaviour of springs in favour of forcing critically damped springs to model the elastic materials.
}

Then we applied Euler only once to solve for $p$:
\begin{equation}
 \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
\end{equation}
with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.  We acknowledge that this is not the proper Newtonian model, however the simulations obtained were closer to the observed behaviour, because there is no inertia in the oscillation.
When experimenting with different integration steps, force values were interpolated between measurements.

\subsection{Collision Detection}
In the simulation, rigid objects (like the robotic finger and the table) are treated as geometric obstacles.  When the simulation of the mesh deformation runs---used to model the sponge---the predicted positions of the vertices may indicate that the mesh penetrates an obstacle.  This should be interpreted as a collision between the sponge and an obstacle.  Since the sponge can not penetrate an obstacle, these overlaps must be eliminated. %The physical consequences of this will affect the system predictions.  For example: if the vertices of the mesh overlap the table, the sponge will remain on top of the table, with its material being more compact, which will cause the response force propagated towards the finger to be increased.

In order to model this, we model the finger and table using geometric primitives.  For the table obstacle, any interpenetration by the sponge is solved by pushing vertices of the mesh of the sponge back to the border of the table.  The springs will then propagate the effect to their neighbours.  For the finger obstacle, resolution is in two stages: by vertex and then by edge.  A vertex will be pushed out to a distance $\epsilon$ of the finger, in the direction of the radius.  For edges, the standard algorithm in \alref{alg:point_segment} is used to estimate the shortest distance between the centre of the finger and the edge.  Given this distance, the algorithm in \alref{alg:push_edge} indicates how far and in which direction the edge must be displaced.

@@

\subsection{Additional Geometric Constraints}
While the mass-spring model incorporates constraints based on mesh geometry, these are soft constraints implemented in the form of an energy function. Even with several of these terms incorporated, such as a term for preservation of angles, the mass-spring model can cause triangles of the mesh to overlap one another and be reversed or flattened.  \comment{All of these situations are innadecuate since this mesh is used to model one face of a 3D object.}  It is thus necessary to employ additional hard geometric constraints to make the mesh internally consistent.  We achieve this through geometric checking subroutines.  There are two geometric constraints that the mesh is enforced to maintain:

\begin{enumerate}
 \item For a reversed triangle, (one vertex crossed over an edge): the vertex gets pushed more than half the distance between the vertex and the edge in the direction of the ``reversed'' height\footnote{$3/4$ of this distance was used in our implementation.}, while both ends of the edge get pushed the same amount in the opposite direction.  This is enough for most cases, but triangles in the border of the mesh can still overlap each other \comment{which would render a bad numerical approximation to the behaviour of a surface}.

@@

The next stage is the learning process which consists in identifying sets of parameters that produce simulations similar to the ground truth.  This problem is addressed in the following section.

\section{Learning to Predict Deformations}
\label{sec:learning}

@@

The contour of the deformable object is used to obtain the ground truth for the training phase, which is the area enclosed by the contour. Since we only require a tracking algorithm which is good enough to evaluate the predictions of the mass-spring system, a very simple one was used.  Geometrically, the contour is represented by a polygon with hundreds of sides, which can also be called a \textit{linear spline} \cite{Menet1991}. The vertices of the polygon work as control points which can be moved to adjust the polygon to the shape of the object as it gets deformed.  Even though this implementation is inspired by the theory of active contours, it works with a much simpler and faster, but less accurate, algorithm, instead of minimizing a global energy.  \comment{This causes some shadowed corners of the material to be missed at some frames, but as \fref{fig:scene}(b) at its bottom left shows, these errors are not significant in comparison to the errors made by the mass-spring model.} This is explained below.  The tracking of the deformable target in the ground truth has two stages:
\begin{enumerate}
 \item The target object is segmented from the environment.  The Canny algorithm is used to extract the edges of the image.  Since the object of interest has a distinctive colour, \comment{its characteristic hue range is extracted from the area enclosed by the initial contour.  Then} the hue of the pixels on every edge are used to keep only edges that could belong to it.  If the average hue of the pixels on an edge falls outside the calibrated range, the edge is eliminated.
 \item The internal representation of its contour, the linear snake, is adapted to its new shape.  The algorithm assumes that no occlusions take place and that the contour does not bend over itself.
\end{enumerate}
For the first frame, the snake is initialised as a rectangle around the target. This rectangle fragments itself into smaller pieces. Each new vertex adheres itself to the closest point on an edge, as reached by growing rings of pixels around its current position, up to a certain threshold distance [\fref{fig:scene}(a)].  If there is no edge within that area, the control point remains where it is. This criterion was selected to make the representation more robust against edges that are easier to detect on some frames than in others.  Once the edge reappears, the control point adapts itself again.  Also, the distance between two consecutive control points is kept within the interval $[min\_distance, max\_distance]$ and two adjacent edges do not form angles smaller than $min\_angle$.  See \alref{alg:regularise}.  This allows the snake to adapt very quickly to the complexity of the contour it represents.

@@

\subsection{Evolutionary Algorithm}
\label{sub:EvolAlg}

Since a mass-spring model is not a constitutive model, its parameters are not directly derived from the physical properties of the material, they instead depend on the geometry of the mesh used to model the object. There can thus be several sets of parameters in different regions of the parameter space that produce adequate simulations \cite{Morris2008}. We therefore chose to use an evolutionary algorithm to both perform global search, and local improvement.

The evolutionary algorithm described in \alref{alg:genetic} was used to search the parameter space of the mass-spring model presented in the previous section.
From the parameters that must be chosen for the simulation to work, some were fixed while others had to be searched for.  They are:
\begin{itemize}
 \item The mass per vertex.  \textit{(For the moment we consider a unitary mass for all vertices.)}
 \item The elasticity constants: for the preservation of length $k_L$, area $k_A$, angles $k_{\varphi}$ and linear damping $k_d$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The plastic parameters: $yield$, $creep$ and the maximum amount of permanent deformation $max\_\alpha$.  \textit{(All springs will have the same values to favour generalisation.)}
 \item The integration time step $\Delta t$.  \textit{(It has been set to 0.1 seconds, even though values as small as 0.01 were tried as well; \comment{however, since they didn't improve stability or accuracy in a visible way, but made the process much slower, they were disregarded as impractical}.)}
 \item An upper limit for the magnitude of the total estimated forces per vertex $max\_Force$.
 \item The minimum area before a triangle is considered \textit{flat}.  (It has been set to $2.0$ units.)
\end{itemize}
To evaluate the suitability of each set of parameters, the simulation is compared with the ground truth tracked from real objects.

\begin{figure}
  \begin{algorithmicieee}{Evolutionary Search}\label{alg:genetic}
 \STATE Let $N$ be the number of sets of parameters to be evaluated per generation and $M$ the number of generations.
 \STATE Let $\alpha + \beta + \gamma + \zeta = N$ denote four portions of the $N$ elements, with: \\
 $\alpha \leftarrow$ number of elements that survive for the next generation as they are, \\
 $\beta \leftarrow$ mutated individuals, \\
 $\gamma \leftarrow$ crossed over, and \\
 $\zeta \leftarrow$ new individuals. \\
 which will be chosen as follows:
 \STATE Let $\sigma_{max} > \sigma_{min}$ both be standard deviations of a Gaussian
 \STATE Let $\Delta\sigma \gets (\sigma_{max} - \sigma_{min})/M$.
 \FOR {$i=1 \to M$}
 \STATE Run $N$ simulations and compare prediction with ground truth (real object), by comparing the area occupied by the real object with the area occupied by the simulated mesh [\fref{fig:scene}(b)].
 \STATE Keep the best $\alpha$ from all simulations.
 \STATE $\sigma \gets \sigma - \Delta\sigma$
 \STATE Generate $\beta$ by adding random Gaussian noise, with standard deviation $\sigma$, to the best candidates.
 \STATE Generate $\gamma$ by selecting values of the parameters from two different sets at random, after eliminating the $\epsilon$ worst.
 \FOR {$\zeta$ elements}
   \STATE Generate completely new sample with:
   \FOR {every parameter}
     \STATE $n \leftarrow randomNumber \in [0,1)$
     \STATE $min \leftarrow \log(minValue)$
     \STATE $max \leftarrow \log(maxValue)$
     \STATE $param \leftarrow e^{(n (max - min) + min)}$
   \ENDFOR
 \ENDFOR
 \STATE $i \gets i + 1$
 \ENDFOR
\end{algorithmicieee}
 \caption{Evolutionary Search for Automatic Calibration}
\end{figure}
